---
title: "Report Preparation"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

# Data

## Loading

First of all, load the raw data. There are 95839 samples and 20 features (including the response variable) here.

```{r}
df <- read.csv("../Data/patient.csv")
dim(df)
head(df)
```

## Cleaning

Aiming to solve the binary classification problem, binary `y` should be elicited first. As for features, not all of them are available. For example, `pneumonia = 99` indicate the feature `pneumonia` is not available for this sample. In this case, we should label `99` as `NA` to avoid future mistakes.

```{r}
df <- df %>% mutate(y = factor(ifelse(death_date == "9999-99-99", 0, 1)),
                    age = factor(age),
              across(c(sex, patient_type, intubated, pneumonia, pregnant, 
                       diabetes, copd, asthma, immunosuppression, hypertension,
                       other_diseases, cardiovascular, obesity, chronic_kidney_failure,
                       smoker, another_case, icu, outcome), ~ factor(ifelse(.>2, NA, .)))) %>%
  select(-c(death_date))
head(df)
```

# Model

In this part, we managed to reproduce the result in the papar \![Classification of Covid-19 Dataset with Some Machine Learning Methods\](<https://dergipark.org.tr/en/pub/jauist/issue/55760/748667>). machine learning classifier methods are considered:

-   BayesNet

-   NaiveBayes

-   SMO (SVM)

-   Random Forest

-   J48 (Tree)

-   IBK (KNN)

## BayesNet

```{r}
library(bnclassify)
library(caret)
```

```{r}
nb <- nb('y', df)
nb <- lp(nb, df, smooth = 0.01)
```



```{r}
res <- cv(nb, df, k = 10, mean = FALSE)
p <- predict(nb, df)
```


```{r}
prop.table(table(p, df$y))
```

```{r}
confusionMatrix(p, df$y)
```













